### 2022-02-19 paper/output/document/thesis_4_applications.tex
started chapter 4
i wrote this chapter in a day lol

%
\begin{isabellebody}%
\setisabellecontext{application}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I 
demonstrate my system's ability to formalize longer, more complicated ethical arguments and present 
the additional capabilities necessary to use my system in practice. In Chapter 2
and Chapter 3, I performed metaethical reasoning while testing different formalizations of the FUL. Metaethical
reasoning analyzes properties of moral thought itself and involves questions about the nature of ethical
truth. This kind of reasoning contrasts with ``applied 
ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. Applied ethical reasoning is relevant with respect to an agent's
particular situation, not just to an ethical theory as an abstract entity. In previous chapters' application tests,
I performed some toy examples of applied ethical reasoning, but this chapter is an extended exploration of
my system's ability to perform applied ethical reasoning. Metaethical reasoning is most interesting to philosophers 
who are trying to formulate the ``best" theory of ethics. Applied ethical reasoning, on the other hand, 
is useful to ordinary people who are trying to decide how to live their lives. In order for automated 
ethics to guide AI agents, it must perform applied ethical reasoning.

One challenge of applied ethical reasoning is that it requires more common sense knowledge than metaethical
reasoning. Because metaethics is about ethics itself, and not about the dilemmas that ethics is 
supposed to help us resolve, this kind of reasoning requires very little knowledge about the world. 
In previous chapters, I perform metaethical reasoning using my system (which formalizes an 
ethical theory) and basic logical facts and objects. Applied ethical reasoning, on the other hand,
focuses on a particular ethical dilemma and thus requires enough 
``common sense" to understand the dilemma and options at hand. For example, an applied ethicist 
evaluating the permissibility of lying needs some robust definition of the term lying and likely some
understanding about the activities of communication and truth telling. Kantians specifically describe
this common sense as ``postulates of rationality" that are nontrivial and nonnormative, but still
part of the process of practical reasoning itself \citep{silber}. \citet{powers} notes that common sense 
is a hurdle for an automated Kantian ethical agent. In this chapter, I attempt to tackle this challenge and
endow my system with this kind of common sense in the specific case of lying. In order for my 
system to be used to perform applied ethical reasoning, it will need to be equipped with a database
of common sense facts and definitions as I present in this chapter. My system will contribute the core
reasoning about the Formula of Universal Law, but this reasoning must be applied to objects that are defined
in this common sense database. In this chapter, I try to understand what this common sense reasoning looks like
for the specific example of lying.

Because these common sense facts can determine my system's judgements, they are part of the trusted
code base for my system. In order for someone to trust my system's judgements, they must trust that the
common sense database is correct becase changing these common sense facts will change the judgements 
that my system makes. For example, if we define truth telling as an act that is self-contradictory (perhaps
by defining it as $p \wedge \neg p$), then my system will output that truth telling is prohibited.
Malicious common sense facts and definitions will result in bad ethical judgements. In other words, garbage in, 
garbage out. This common sense reasoning is also not a part of the system I contribute here—in order for my
system to be used, future work must develop a common sense database for the particular situation at hand. 
The challenge of endowing automated ethical reasoners with common sense reasoning is not unique to my 
system, and virtually all prior attempts in machine ethics face similar challenges.\footnote{See Section
Related Work for a survey of the common sense required in prior work.} Most prior attempts
sidestep this question, whereas I contribute an prototype implementation of one kind of common sense reasoning.

The specific kind of common sense reasoning required appears to be a challenge 
for automating Kantian ethics, and may imply that consequentialist or virtue ethical automated
agents are within closer reach for automation. Ultimately, Kantian ethics is still
easiest to automate because it will require fewer, less controversial common sense facts than other ethical
systems. As the examples in this section demonstrate, Kantian ethics requires a definition of lying (which any other 
theory would also requires) and the knowledge that if everyone lies in a given context, no one will believe 
each other in that particular context. This latter fact may not be required for every ethical theory, 
but is relatively uncontroversial. It is a kind of intuition about human behavior that is generally accepted. 
Neither a definition of lying nor this property of lying seem like unreasonable prerequisites for ethical reasoning.

Consequentialism, on the other hand, would require much more specific data about the consequences of 
a lie, perhaps for the specific person's credibility, for the victim of the lie, for the third-parties 
watching the lie unfold. Consequentialism requires more numerous and specific judgements, all of which are 
likely to be more controversial than the two outlined above for Kantian ethics. Similarly, virtue ethics
would likely require information about the actor's entire moral character, including their attitude
towards the lie and their moral history. Virtue ethics would also require robust definitions of the 
relevant virtues, in addition to a definition of lying. Thus, while Kantian ethics requires some 
common sense reasoning, it requires fewer and less controversial background facts than other ethical theories.

The extended examples presented in this section also demonstrate the difficulty of formulating a 
maxim to pass as input to my implementation of the FUL. A large part of the challenge of applied Kantian
ethics is formulating a maxim that accurately captures an agent's principle of action, so a totally
automated agent using my system will need some way to formulate these maxims well. In this section,
I manually implement Korsgaard's formulation of certain maxims, and I will later argue that manual formulation
of maxims is, at present, the way forward.

This chapter will provide additional examples of the kinds of common sense facts and maxim formulation required to get my system
off the ground. I will aim to use a lean and uncontroversial common sense database
to achieve robust and powerful results. This serves as evidence for the ease of automating
Kantian ethics, an example of what additional work my system requires, and a demonstration of the contributions
that I make. These examples demonstrate that nuanced common sense facts and maxims can cause my system to contribute 
nuanced judgements faithful to philosophical literature. On one hand, this means that my system can perform
sophisticated ethical reasoning, but on the other, the quality of this reasoning is heavily reliant on 
the common sense database and the input maxim. Thus, my system \emph{could} make smart ethical judgements 
in practice, but getting to that point will require a robust, trusted common sense database and maxim
formulator.

In this paper, I contribute an implementation of the Formula of Universal Law that, when equipped with 
relatively thin common sense facts, can performed nuanced, sophisticated ethical reasoning. When given 
as input a maxim in the format specified in Chapter 3, my system can show that the maxim is permissible, 
obligatory, or prohibited. It can provide a verifiable proof of its judgement and a human-readable list
of the facts or axioms it used to reach its conclusion. My system is faithful to philosophical literatire
on Kantian ethics and serves as the first step towards a fully or partially automated ethical AI agent. 
It also establishes the power and potential of computational ethics, or the use of computational tools
for ethical investigation, as explored in this section's Applied Ethical reasoning.

\emph{Stick a bit here about the philosophical work that will go in this section}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Simple Lying Examples \label{sec:lyingprelim}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
This chapter focuses on the example of lying because this case is hotly debated in the Kantian
literature. I draw on examples of ethical reasoning as presented in Korsgaard's ``Right to Lie," which 
examines exactly how strict Kant's prohibition on lying is. She picks up a long-running debate in the 
literature through the example of someone who shows up at your door and asks, ``Is Sara home?" Unbeknownst to them, 
you know that they want to know Sara's location in order to murder her. Ordinary moral intuition 
prescribes that, if Sara is home, you should lie and say that she is not in order to 
protect her from the murderer, but the categorical imperative seems to prohibit lying in all 
circumstances. In this section, I will formalize Korsgaard's treatment of lying and joking under the 
categorical imperative, focusing on the common sense assumptions necessary to achieve her conclusions. 
In the next section, I will formalize the core of Korsgaard's argument that the categorical imperative
coheres with ordinary intuition and does not prohibit lying to the murderer to protect Sara.

First, Korsgaard argues that the categorical imperative appears to prohibit all lies because, when universalized, 
lies will no longer be believed. Thus, lying could never be an effective way of achieving any goal when universalized.
Korsgaard points out that ``we believe what is said to us in a given context because most of the time 
people in that context say what they really think" \citep[4]{KorsgaardRTL}. In order to formalize this argument,
I first need to define the relevant terms.%
\ believe{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\ believes\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\isamarkupcmt{Person s::subject believes sentence t::term. The concept of belief will play a crucial role both in 
the arguments for lying being prohibited and for the maxim of lying to the murderer being permissible. 
Logicians and epistemologists have developed robust, complex logics of belief and knowledge \citep{seplogicbelief}. For the 
sake of this project, I avoid this complexity by merely defining the concept of belief as an empty
constant that maps a subject, term pair to a term. For the examples in this section, this choice suffices, 
as my common sense beliefs encode enough properties of belief for my purposes. Future work could
integrate a much more complex logic of belief into my system.%
\isanewline
\ utter{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\isanewline
\isamarkupcmt{Person s::subject utters sentence t::term.%
\ utter{\isacharunderscore}falsehood{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}utter{\isacharunderscore}falsehood\ s\ t\ {\isasymequiv}\ \ {\isacharparenleft}utter\ s\ t{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ t{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{Person s utters falsehood t if and only if s utters t and t is false.%
\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t\ {\isasymequiv}\ {\isacharparenleft}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}believe\ s\ t{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{Person s::subject knowingly utters a falsehood t, if they both utter t as a falsehood and don't
believe t. This and the above abbreviations are the core of my formalization of Korsgaard's definition
of lying. They are also relatively uncontroversial and have little normative content.%
\isanewline
\ lie{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}lie\ {\isasymequiv}\ {\isasymlambda}\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}{\isachardot}\ {\isasymexists}t{\isachardot}\ {\isacharparenleft}a\ \isactrlbold {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymlambda}s{\isachardot}\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymexists}p{\isachardot}\ {\isasymforall}w{\isachardot}\ {\isacharparenleft}g\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}believe\ p\ t{\isacharparenright}{\isacharparenright}\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{Using the above definitions, I can characterize a maxim as a lie if 
    (a) the act requires knowingly uttering a falsehood and,
    (b) the end requires that some person $p$ believe the false statement $t$.%
To avoid unintentional wrongdoing, I focus on ``knowing lies," 
in which the speaker is aware that they are lying. It is uncontroversial that, in order for an act to be
a knowing lie, the speaker must utter a false statement that they do not believe. The second half of 
this definition requires that the goal of the lie is deception and is inspired by  Korsgaard's interpretation 
of a lie. She understands a lie as a kind of falsehood that is usually effective \emph{because} it decieves
\citep[4]{KorsgaardRTL}. In my formalization, this means that the purpose or goal of the maxim must
involve decieving someone, or, in other words, that someone believe the falsehood told.%
With the above logical background, I can now automate Korsgaard's argument that maxims that involve

@@

\isamarkupcmt{I am trying to reason about \emph{my} obligations so I will define myself as a specific subject.%

@@

this lemma represent the logical and common sense background necessary to reach this conclusion. This

@@

relatively ``thin" premises about individual behavior, my system can derive a prohibition against lying.%

@@

Korsgaard's core piece of ``common sense" about lying \cite[5]{KorsgaardRTL}. This simple assumption encodes the common sense knowledge

@@

lying" fact. In the rest of this section, I will test versions of this assumption, effectively encoding 

@@

\isamarkupcmt{Restrict our focus to worlds in which the circumstances hold, as these are the morally interesting 
worlds for this example. A technical detail.%

@@

\isamarkupcmt{Unlike many of the other proofs in this project, this proof is a little heavier and requires some manual
work to produce. After I divide the proof into the intermediate steps shown here, and Isabelle is able to do 

@@

which is quite close to the idea of the maxim not being universalizable.%

@@

$\vDash prohibited\! m \! me$.%

@@

Now that I have formalized Korsgaard's argument for why lying is prohibited, I will 
implement her argument for why jokes are permissible. Specifically, she defines a joke as a story that is 
false and argues that joking is permissible because ``the universal practice of lying in the context of jokes
does not interfere with the purpose of jokes, which is to amuse and does not depend on
deception" \citep[4]{KorsgaardRTL}. First, I define a joke.%

@@

\ joke{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}joke\ {\isasymequiv}\ {\isasymlambda}\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}{\isachardot}\ \ {\isasymexists}t{\isachardot}\ {\isacharparenleft}a\ \isactrlbold {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymlambda}s{\isachardot}\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}\ {\isacharparenleft}{\isasymexists}p{\isachardot}\ {\isasymforall}w{\isachardot}\ {\isacharparenleft}g\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}believe\ p\ t{\isacharparenright}{\isacharparenright}\ w{\isacharparenright}{\isachardoublequoteclose}%

@@

This definition of a joke merely defines a joke as a falsehood uttered for some purpose that 
doesn't require deception, where deception involves someone believing the uttered falsehood. Notice 
that this is quite a thin definition of a joke; it doesn't require any conception of humor but merely
distinguishes jokes from lies. As far as common sense reasoning goes, this is a relatively tame proposition. 
I will now demonstrate that this conception of a joke is sufficient to show that joking is permissible.

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}applications}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I use my system to produce sophisticated, nuanced
responses to two moral dilemmas. These dilemmas serve as examples of how my system could be used
in practice and demonstrate my system's ability to formalize longer, more complicted ethical arguments
than those presented in Chapter 3. 
Many of the tests in Section \ref{testing} perform metaethical reasoning, which analyzes properties
of morality itself and involves questions about the nature of ethical truth. In this chapter, I perform
``applied ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. This is the kind of reasoning necessary for an AI agent using my system to
navigate the real world.

One challenge of applied ethical reasoning is that it requires more factual background than metaethical
reasoning. Because metaethics is about ethics itself, and not about the dilemmas that ethics is 
supposed to help us resolve, this kind of reasoning requires very little knowledge about the world. 
Applied ethical reasoning, on the other hand,
focuses on a particular ethical dilemma and thus requires enough factual background, or common sense, 
to understand the dilemma and options at hand. For example, an applied ethicist 
evaluating the permissibility of lying needs a robust definition of the term lying and likely some
understanding of communication and truth telling. Kantians specifically describe
this common sense as ``postulates of rationality'' that are nontrivial and nonnormative, but still
part of the process of practical reasoning itself \citep{silber}.

In this chapter, I attempt to tackle this challenge by
endowing my system with this kind of common sense in the specific case of lying. In order for my 
system to perform applied ethical reasoning, it will need common sense facts and definitions, like
those I present in this chapter. My system contributes the ability to reason
using the Formula of Universal Law, but this reasoning must be applied to objects that are defined
using common sense. In this chapter, I describe such common sense for the specific example
of lying.

Because these common sense facts can determine my system's judgements, they are part of the trusted
code base for my system, or the logic and code that a user must trust in order to trust my system. 
Changing these common sense facts will change the judgements 
that my system makes. For example, if we define truth telling as an act that is self-contradictory (perhaps
by defining it as $p \wedge \neg p$), then my system will output that truth telling is prohibited.
Malicious common sense facts and definitions will result in bad ethical judgements. 
The challenge of endowing automated ethical reasoners with common sense reasoning is not unique to my 
system, and virtually all prior attempts in machine ethics face similar challenges.\footnote{See Section
Related Work for a survey of the common sense required in prior work.} Many prior attempts
sidestep this question, whereas I contribute an prototype implementation of one kind of common sense reasoning.

This chapter will provide examples of the kinds of common sense facts required to get my system
off the ground. I use a lean and uncontroversial common sense database
to achieve robust and powerful results. This serves as evidence for the ease of automating
Kantian ethics, an example of the additional work required to use my system in practice, and a 
demonstration of my system's power and flexibility. These examples demonstrate that, armed with nuanced 
common sense facts, my system can make sophisticated judgements faithful to philosophical literature.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Lies and Jokes \label{joking}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The moral status of lying is hotly debated in the Kantian literature. I focus on two dilemmas
presented in Korsgaard's ``Right to Lie," which 
examines how strict Kant's prohibition on lying is \citep{KorsgaardRTL}. She begins with the case of 
lying and joking. Many of Kant's critics argue that his prohibition on lies includes lies told in 
the context of a joke, which should be permissible. Korsgaard responds by arguing 
that there is a crucial difference between lying and joking: lies involve deception, but jokes do not. 
The purpose of a joke is amusement, which does not rely on the listener believing the story told. 
Given appropriate definitions of lies and jokes, my system shows that jokes are permissible but lies 
are not, demonstrating its power and flexibility. This section demonstrates how my system can be used 
in practice; it needs to be given some baseline common sense facts, but with those facts, it can make 
sophisticated judgements. Moreover, because my system is faithful to definitions found in philosophical 
literature, it can perform nuanced reasoning, demonstrating the value of faithful automated ethics.

First, Korsgaard argues that the categorical imperative prohibits lies because they involve deception. 
When universalized, lies will no longer be believed, so lying could never be an effective way of achieving 
any goal when universalized. Korsgaard points out that ``we believe what is said to us in a given 
context because most of the time people in that context say what they really think'' \citep[4]{KorsgaardRTL}. 
In order to formalize this argument, I first need to define the relevant terms and assumptions, which include
lying and Korsgaard's argument about the basis of trust.

I define lying and trust in terms of belief. As in Section \ref{details}, I choose thin, or minimal,
definitions of terms like belief to reduce the potential for controversy in my system's factual background.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ believe{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\ believes\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isamarkupcmt{\texttt{believe} is a constant that maps a subject and a term to another DDL term. For example, 
subject ``Sara'' might believe the term ``the sky is blue'' to create the sentence ``Sara believes
that the sky is blue,'' which can be true or false at a world.%
}%
\begin{isamarkuptext}%
Logicians and epistemologists develop and debate complex logics of belief and knowledge \citep{seplogicbelief}. 
I avoid this complexity by defining the concept of belief simply as a constant that maps a subject, term pair 
to a term. For the examples in this section, this choice suffices. 
I encode some minimal properties of belief below, but avoid any full definition of the term.

Belief is useful to construct the idea of ``knowingly uttering a falsehood,'' a core component 
of both lying and joking.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ utter{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{\texttt{utter} also maps a subject and term to another DDL term. For example, the sentence ``Sara
utters, `I am hungry' '' is a DDL term that can be true or false at a world.%
}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ utter{\isacharunderscore}falsehood{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}utter{\isacharunderscore}falsehood\ s\ t\ {\isasymequiv}\ \ {\isacharparenleft}utter\ s\ t{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ t{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{To utter a falsehood is to utter a statement that is false, or to utter $t$ when $\neg t$.%
}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}s{\isasymRightarrow}t{\isasymRightarrow}t{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t\ {\isasymequiv}\ {\isacharparenleft}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}believe\ s\ t{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Sometimes we unknowingly utter falsehoods. For example, if I believe that the Earth is flat, then 
when I utter, ``the Earth is flat,'' I am unknowingly uttering a falsehood. This motivates defining
the idea of knowingly uttering a falsehood, which requires both uttering a falsehood and not believing 
your utterance. If I utter ``the Earth is flat,'' even though I know that the Earth is round, I am 
knowingly uttering a falsehood.%
}%
\begin{isamarkuptext}%
The above abbreviations are the core of my formalization of Korsgaard's definitions
of lying and joking. They are also relatively uncontroversial and encode little moral
or normative content. They say nothing about the moral status of uttering a falsehood, 
the agent's intention when making the utterance, or the conversational norms guiding the utterance. 
The complexities of a complete definition of lying or belief are 
unnecessary for Kantian ethics, and therefore for my system, to make moral judgements.

Using the above definitions, I define lying. I characterize a maxim as involving a lie if 
the act requires knowingly uttering a falsehood and the end requires that some person $p$ believe 
the false statement $t$.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lie{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}lie\ {\isasymequiv}\ {\isasymlambda}\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}{\isachardot}\ {\isasymexists}t{\isachardot}\ {\isacharparenleft}a\ \isactrlbold {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymlambda}s{\isachardot}\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymexists}p{\isachardot}\ {\isasymforall}w{\isachardot}\ {\isacharparenleft}g\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}believe\ p\ t{\isacharparenright}{\isacharparenright}\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{The abbreviation above maps a maxim to a boolean value that indicates if it is a lie.%
}%
\begin{isamarkuptext}%
To avoid unintentional wrongdoing, I focus on ``knowing lies,'' 
in which the speaker is aware that they are lying. It is uncontroversial that, in order for an act to be
a knowing lie, the speaker must utter a false statement that they do not believe. This also makes it 
easier to make moral judgements about the speaker's action, since they were, at the very least, aware
of their lie. The second half of
this definition requires that the goal of the lie is deception. This is inspired by  Korsgaard's interpretation 
of a lie. She understands a lie as a kind of falsehood that is usually effective \emph{because} it decieves
\citep[4]{KorsgaardRTL}. In my formalization, this means that the purpose or goal of the maxim must
involve decieving someone, or, in other words, that someone believe what the speaker knows to be a 
falsehood.

With the above logical background, I automate Korsgaard's argument that maxims that involve

@@

\isamarkupcmt{I am trying to reason about \emph{my} obligations so I will define myself as a specific subject. Again,
this is a minimal definition that does not include any facts about me, such as the fact that my name is
Lavanya or that I have brown hair.%

@@

this lemma represent the common sense necessary to reach this conclusion. This

@@

relatively minimal premises about individual behavior, my system derives a prohibition against lying.%

@@

Korsgaard's core piece of ``common sense'' about lying \cite[5]{KorsgaardRTL}. This simple assumption encodes the common sense knowledge

@@

trust'' fact. In the rest of this section, I will test versions of this assumption, effectively encoding 

@@

\isamarkupcmt{Restrict our focus to worlds in which the circumstances hold. A technical detail.%

@@

\isamarkupcmt{This proof requires some manual work. After I divide the proof into the intermediate steps shown 
here, Isabelle is able to do 

@@

which is close to the idea of the maxim not being universalizable.%

@@

$\vDash prohibited \; m \; me$.%

@@

The lemma above demonstrates that my system finds that lying is prohibited with a thin definition
of lying and relatively uncontroversial facts about the world. The logical background needed is the fact that lying
requires knowingly uttering a falsehood with the goal that someone believe the falsehood, a definition
of lying that is relatively well-accepted. The factual background needed consists of the fact that if 
everyone lies in a given context, then people will stop believing each other in that context. This is 
a slightly heavier assumption, but it is still so uncontroversial that Korsgaard doesn't bother to justify
it in her argument against lying \citep{KorsgaardRTL}.

Now that I have formalized Korsgaard's argument for why lying is prohibited, I will
implement her argument for why jokes are permissible. Specifically, she defines a joke as a story that is 
false and argues that joking is permissible because ``the universal practice of lying in the context of jokes
does not interfere with the purpose of jokes, which is to amuse and does not depend on
deception'' \citep[4]{KorsgaardRTL}. I use the fact that a joke does not depend on deception as the 
defining feature of a joke.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ joke{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}joke\ {\isasymequiv}\ {\isasymlambda}\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}{\isachardot}\ \ {\isasymexists}t{\isachardot}\ {\isacharparenleft}a\ \isactrlbold {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymlambda}s{\isachardot}\ knowingly{\isacharunderscore}utter{\isacharunderscore}falsehood\ s\ t{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}\ {\isacharparenleft}{\isasymexists}p{\isachardot}\ {\isasymforall}w{\isachardot}\ {\isacharparenleft}g\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}believe\ p\ t{\isacharparenright}{\isacharparenright}\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Thie abbreviation states that a maxim is a joke if the action involves knowingly uttering a falsehood but
the goal does \emph{not} require that someone believe the falsehood told.%
}%
\begin{isamarkuptext}%
This definition of a joke defines a joke as a falsehood uttered for some purpose that 
doesn't require deception, where deception involves someone believing the uttered falsehood. 
This definition doesn't require any conception of humor, but merely
distinguishes jokes from lies.

Korsgaard argues that her above argument for a prohibition against lying also implies that joking is 
permissible, because its purpose is not to decieve, but something else entirely. This means that, 
even armed with the same convention of trust assumption as above, joking should be permissible. The 
lemma below shows exactly that.%

@@

\ joking{\isacharunderscore}not{\isacharunderscore}prohibited{\isacharcolon}\isanewline
\isamarkupcmt{Initial set-up: $m$ is a well-formed maxim composed of some circumstances, act, and goal.%
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}joke\ m{\isachardoublequoteclose}\isanewline
\isamarkupcmt{$m$ is a maxim about joking. Precisely, it is a maxim in which the action is to knowingly utter a 
falsehood and the goal does not require that someone believe this falsehood.%
\isamarkupcmt{The same convention of trust assumption as in the above example.%
\isamarkupcmt{Restrict our focus to worlds in which the circumstances hold. A technical detail.%
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymTurnstile}\ {\isacharparenleft}permissible\ m\ me{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}smt\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ case{\isacharunderscore}prod{\isacharunderscore}conv{\isacharparenright}\isanewline
\isamarkupcmt{Isabelle is able to show that maxims about joking are permissible. It also lists the facts used 
in its proof, which offer insight into how it arrived at its judgement. Specifically, it uses assumptions 1, 
2, and 3, which are the logical background and definition of the joking maxim. Notably, it does not use the 
convention of trust assumption. This demonstrates that even the convention of trust assumption is not
strong enough to prohibit joking, which is exactly the desired result.%

@@

My system shows that lies are prohibited and jokes are permissible with thin 
conceptions of amusement and deception. This shows that it isolates a necessary and sufficient property 
of this class of maxims that fail the universalizability test. My definitions of a lie and joke
only differ in whether or not their goal requires that someone believe the falsehood in question, so
this is a necessary and sufficient condition for a maxim about knowingly uttering falsehoods to be prohibited.
This logical fact derived by my system tracks a fact implicit in Korsgaard's argument and in most Kantian accounts
of lying: the wrongness of lying is derived from the requirement that someone believe the falsehood. 
The logical reality that this property is necessary and sufficient to generate a prohibition reflects
a deep philosohopical explanation of \emph{why} certain maxims about uttering falsehood fail the universalizability
test. Universalizing uttering a falsehood makes belief in that 
falsehood impossible, so any maxims with goals that require believing in the falsehood will be prohibited.

This account not only describes the kind of maxims that fail or pass the universalizability test, but it 
also provides a guide to constructing permissible maxims about uttering falsehoods. As an example, 
consider the idea of throwing a surprise birthday party. At first glance, the maxim of action is 
something like, ``When it is my friend's birthday, I will secretly plan a party so that I can surprise
them.'' The goal ``so that I can surprise them'' clearly requires that your friend believe the falsehood that 
you are not planning a party, else the surprise would be ruined. This seems to imply that 
Kantian ethics would prohibit surprise parties, which is a sad conclusion for birthday-lovers everywhere. 
Knowing that this maxim is prohibited because the goal requires belief
in a falsehood provides a way to rescue surprise parties. When throwing a 
surprise party, the objective is \emph{not} to surprise your friend, but to celebrate
your friend and help them have a fun birthday. If someone ruins the surprise, but the party is still fun
and the birthday person feels loved, then such a party is a success! Someone who
calls this party a failure is clearly missing the point of a surprise party. The goal of 
a surprise party is not the surprise itself, but rather celebrating the birthday person. The modified
goal no longer requires belief in the falsehood and thus passes the universalizability test.
